PanGraph: scalable bacterial pan-genome graph construction

### 2022-09-13 main.tex
feat: continue supplementary
feat: polished increasing size figure

\paragraph{Merging of homologous sequence.}
\emph{Pancontigs} encapsulate linear multiple-sequence alignments which are modelled internally by a star phylogeny, i.e. are assumed to be well-described by a reference sequence augmented by independent SNPs and indels for each contained isolate.
Importantly, during the all-to-all alignment phase, \emph{pancontigs} are aligned based \emph{solely} upon their consensus sequence.
All putative homologous alignments found by \emph{pancontig} alignment are ranked according to the pseudo-energy
\begin{equation}
    E = -\ell + \alpha N_c + \beta N_m
\end{equation}
where $\ell$, $N_c$, and $N_m$ denote the alignment length, number of \emph{pancontigs} created by the merger, and number of polymorphisms per genome in the newly created {pancontig} respectively.
See Fig \ref{fig:visualization}B for a graphical definition for each term.
Additionally, $\alpha$ and $\beta$ are hyperparameters of the algorithm, respectively controlling the tradeoff between fragmentation of the graph and the maximum diversity within each block.
Only alignments with negative pseudo-energy are performed.

\begin{figure}[htb]
    \includegraphics[width=.5\textwidth]{algorithm.png}
    \caption{{\bf Overview of PanGraph algorithm}.
    (A) The alignment graph is constructed progressively by aligning graphs pairwise up a guide tree constructed from neighbor-joining the minimizer overlap between strains. 
    (B) During pairwise alignment, \emph{pancontigs} (blue and green)
    are merged by identifying homologous intervals (shown in yellow).
    If the underlying alignments are viewed compatible, i.e. the energy is less than 0, the pancontigs are merged.
    }
    \label{fig:visualization}
\end{figure}

@@

\begin{figure}[htb]
    \includegraphics[width=.4\textwidth]{benchmark.pdf}
    \caption{{\bf Algorithm performance}.
        PanGraph scales linearly with the number of input genomes.
        This is a direct result of the guide tree simplification.
        The solid line and ribbons display the mean and standard deviation respectively.
        All runs were performed on an AMD Ryzen 7 2700X utilizing 8 cores.
    }
    \label{fig:toy-performance}
\end{figure}

@@

\begin{figure*}[htb]
    \includegraphics[width=.8\textwidth]{figs/incr_size.pdf}
    \caption{{\bf Pangenome graph properties vs. dataset size}}
    \label{fig:panx-size}
\end{figure*}

>>>>

\paragraph{Merging of homologous sequence.}
\emph{Pancontigs} encapsulate linear multiple-sequence alignments which are modelled internally by a star phylogeny, i.e. are assumed to be well-described by a reference sequence augmented by independent SNPs and indels for each contained isolate.
Importantly, during the all-to-all alignment phase, \emph{pancontigs} are aligned based \emph{solely} upon their consensus sequence.
All putative homologous alignments found by \emph{pancontig} alignment are ranked according to the pseudo-energy
\begin{equation}\label{eq:pseudo-energy}
    E = -\ell + \alpha N_c + \beta N_m
\end{equation}
where $\ell$, $N_c$, and $N_m$ denote the alignment length, number of \emph{pancontigs} created by the merger, and number of polymorphisms per genome in the newly created {pancontig} respectively.
See Fig \ref{fig:visualization}B for a graphical definition for each term.
Additionally, $\alpha$ and $\beta$ are hyperparameters of the algorithm, respectively controlling the tradeoff between fragmentation of the graph and the maximum diversity within each block.
Only alignments with negative pseudo-energy are performed.

\begin{figure}[htb]
    \includegraphics[width=.45\textwidth]{figs/algorithm.png}
    \caption{{\bf Overview of PanGraph algorithm}.
        (A) The alignment graph is constructed progressively by aligning graphs pairwise up a guide tree constructed from neighbor-joining the minimizer overlap between strains.
        (B) During pairwise alignment, \emph{pancontigs} (blue and green)
        are merged by identifying homologous intervals (shown in yellow).
        If the underlying alignments are viewed compatible, i.e. the energy is less than 0, the pancontigs are merged.
    }
    \label{fig:visualization}
\end{figure}

@@

\begin{figure}[htb]
    \includegraphics[width=.4\textwidth]{figs/benchmark.pdf}
    \caption{{\bf Algorithm performance}.
        PanGraph scales linearly with the number of input genomes.
        This is a direct result of the guide tree simplification.
        The solid line and ribbons display the mean and standard deviation respectively.
        All runs were performed on an AMD Ryzen 7 2700X utilizing 8 cores.
    }
    \label{fig:toy-performance}
\end{figure}

@@

\begin{figure*}[htb]
    \includegraphics[width=.9\textwidth]{figs/incr_size.pdf}
    \caption{{\bf Pangenome graph properties vs. dataset size}. We build pangenome graphs with increasing number of isolates from the E.~Coli dataset and measure the scaling of different properties of the pangenome graph. To quantify variability, we repeat the pangraph construction 10 times for each size (except the complete graph containing all 307 isolates), with different random sets of isolates. Lines and shaded areas represent means and standard deviations over these 10 measurements.
        \textbf{A}: n. of pancontigs in the graph. We count the total number of pacontigs (blue), the number of core pancontigs (orange) and the minimum number of pancontigs that contain more than 50\% of the pangenome (L50, black).
        \textbf{B}: average size of pancontigs (blue), of only core pancontings (orange), and size of the smallest panconting in the minimal set that spans 50\% of the pangenome (N50, black).
        \textbf{C}: cumulative size of all genomes in the pangenome graph (gray), total pangenome size (blue) and size of core pangenome (orange).
    }
    \label{fig:panx-size}
\end{figure*}
